### 2020-08-24 paper/definitions.tex
Corrections

\begin{defn}[connected graph]
  A graph $G$ is \emph{connected} if for every pair $\{x, y\} \subseteq V(G)$ of distinct vertices, there is a path from $x$ to $y$. \todo{formal enough?}
\end{defn}

@@

Sometimes, we will call a complement of a member of a class $\Gamma$ an \emph{anti-
  $\Gamma$}, e.g. graph $G = (\{v_1, v_2, v_3, v_4\}, \{v_1v_3, v_2v_4\})$ is an anticycle.

@@

Therefore an \emph{anticlique} is a graph in which there are no edges. We will also call anticliques \emph{independent sets}. In a similar fashion, given a graph $G = (V, E)$, a subset of its vertices $V' \subseteq V$ will be called \emph{independent} (in the context of $G$) iff $G[V']$ is an anticlique.

@@

\begin{defn}[Berge graph]
  A graph $G$ is \emph{Berge} if both $G$ and $\overline{G}$ have no odd hole.
\end{defn}

>>>>

\begin{defn}[connected graph, subset]
  A graph $G$ is \emph{connected} if for every pair $\{x, y\} \subseteq V(G)$ of distinct vertices, there is a path from $x$ to $y$. \todo{formal enough?}
  A subset $X \subseteq V(G)$ is connected if the graph $G[X]$ is connected.
\end{defn}

@@

% Sometimes, we will call a complement of a member of a class $\Gamma$ an \emph{anti-
% $\Gamma$}, e.g. graph $G = (\{v_1, v_2, v_3, v_4\}, \{v_1v_3, v_2v_4\})$ is an anticycle.

\begin{defn}[anticonnected graph, subset]
  A graph $G$ is anticonnected if $\overline{G}$ is connected.
  A subset $X$ is anticonnected if $\overline{G}[X]$ is connected.
\end{defn}

@@

\begin{defn}[anticlique]
  An \emph{anticlique} is a graph in which there are no edges. We will also call anticliques \emph{independent sets}. 
\end{defn}
In a similar fashion, given a graph $G = (V, E)$, a subset of its vertices $V' \subseteq V$ will be called \emph{independent} (in the context of $G$) iff $G[V']$ is an anticlique.

@@

\begin{defn}[Berge graph]
  \label{def:bergeGraph}
  A graph $G$ is \emph{Berge} if both $G$ and $\overline{G}$ have no odd hole.
\end{defn}

### 2020-08-24 paper/main.tex
Corrections

\documentclass{article}

\usepackage{appendix}
\usepackage{todonotes}
\usepackage{wrapfig}
\usepackage{intcalc}
\usepackage{amsthm, amsfonts, thmtools}
\usepackage{cleveref}
\usepackage{floatflt}

\usepackage{tikz}
\tikzset{%
    simplegraph/.style={every node/.style={draw,circle, inner sep=0pt, minimum size=6mm}},
}
\usetikzlibrary{snakes}
\usetikzlibrary{shapes.geometric}
\usetikzlibrary{positioning, fit, calc}

\usepackage[
	style=numeric
]{biblatex}
\addbibresource{bbl.bib}

\newtheorem{theorem}{Theorem}[section]
\newtheorem{corollary}{Corollary}[theorem]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{defn}{Definition}[section]
\renewcommand{\listtheoremname}{List of definitions}

\newcommand{\TODO}{\todo[inline]}
\newcommand\Lovasz{Lovász }
\newcommand\T{\mathcal{T}}
\crefname{lemma}{Lemma}{Lemmas}

% Language and typesetting notes:
%
% - Color, not Colour (American English)
% - Node not a Vertex of a graph

\author{Adrian Siwiec}
\date{\today{}}
\begin{document}
\begin{titlepage}
	\begin{center}

		\large
		\textbf{Jagiellonian University}\\
		Department of Theoretical Computer Science\\

		\vspace{1.5cm}

		\Large
		\textbf{Adrian Siwiec}

		\vspace*{2cm}

		\textbf{\LARGE Perfect Graph Recognition and Coloring}

		\vspace{0.5cm}
		\large

		\vfill
		\Large
		Master Thesis

		\vfill
		\Large
		Supervisor: dr inż. Krzysztof Turowski

		\vspace{0.8cm}

		September 2020

\end{center}
\end{titlepage}

\begin{abstract}
TODO
\end{abstract}

@@

\section{Definitions}
\input{definitions.tex}

@@

A definition of perfect graphs (\cref{def:perfectGraph}) states that given a graph $G$, the chromatic number and cardinality of the largest clique of \emph{its every induced subgraph} should be equal. The notion of perfect graphs was first introduced by Berge in 1961 \cite{CB61} and it indeed captures the idea of graph being ''simple'' -- in all perfect graphs the coloring problem, maximum clique problem, and maximum independent set problem can be solved in polynomial time \cite{grotschel1993}

\begin{theorem}{Perfect graph theorem}
A perfect graph theorem, first conjured by Berge in 1961~\cite{CB61} and then proven by \Lovasz in 1972~\cite{LL72} states that a graph is perfect iff its complement graph is also perfect. \todo{Should we give some proof of that here? Maybe based on proof in \cite{GC03}}
\end{theorem}

A \emph{hole} is an induced chordless cycle of length at least 4. An \emph{antihole} is an induced subgraph whose complement is a hole. A \emph{Berge} graph is a graph with no holes or antiholes of odd length.

In 1961 Berge conjured that a graph is perfect iff it is Berge in what has become known as a strong perfect graph conjecture. In 2001 Chudnovsky et al. have proven it and published the proof in an over 150 pages long paper \citetitle{MC06} \cite{MC06}. The following overview of the proof will be based on this paper and on an article withe the same name by Cornuéjols \cite{GC03}.

@@

Odd holes are not perfect, since their chromatic number is 3 and their largest cliques are of size 2. It is also easy to see, that an odd antihole of size $n$ has a chromatic number of $\frac{n+1}{2}$ and largest cliques of size $\frac{n-1}{2}$. It is therefore clear, that if a graph is not Berge it is not perfect. To prove that every Berge graph is perfect is the proper part of the strong perfect graph theorem.

\TODO{How long and detailed overview of the proof should we provide?}

@@

A \emph{path} in $G$ is an induced subgraph that is connected, with at least one node, no cycle and no node of degree larger than 2 (sometimes called chordless path). The \emph{length} of a path or a cycle is the number of edges in it. \todo{Should we move these definitions elsewhere?} A \emph{triangle} in a graph is a set of three pairwise adjacent nodes.

@@

\begin{wrapfigure}{r}{0.35\textwidth}
	\centering\begin{tikzpicture}[scale=0.7,simplegraph]
	\node(a) at (0,0) {a};
	\node(b1) at (-2,-6) {$b_1$};
	\node(b2) at (0,-4.268) {$b_2$};
	\node(b3) at (2,-6) {$b_3$};

	\node(P12) at (-2/4, -6/4) {\small$P_{12}$};
	\node(P13) at (-4/4, -12/4) {\small$P_{13}$};
	\node(P14) at (-6/4, -18/4) {\small$P_{14}$};

	\node(P32) at (2/3, -6/3) {\small$P_{32}$};
	\node(P33) at (4/3, -12/3) {\small$P_{33}$};

	\draw (b1) to (b3);
	\draw (b1) to (b2);
	\draw (b2) to (b3);

	\draw (a) to (P12);
	\draw (P12) to (P13);
	\draw (P13) to (P14);
	\draw (P14) to (b1);

	\draw (a) to (P32);
	\draw (P32) to (P33);
	\draw (P33) to (b3);

	\draw (a) to (b2);
\end{tikzpicture}%
\caption{An example  of a pyramid.}%
\vspace{-1.2cm}
% \end{floatingfigure}
\end{wrapfigure}

A \emph{pyramid} in G is an induced subgraph formed by the union of a triangle $\{b_1,b_2,b_3\}$, three paths $\{P_1, P_2, P_3\}$ and another node $a$, so that:
\begin{itemize}
	\item $\forall_{1\leq i \leq 3}$ $P_i$ is a path between $a$ and $b_i$
	\item $\forall_{1\leq i < j \leq 3}$ $a$ is the only node in both $P_i$ and $P_j$ and $b_ib_j$ is the only edge between $V(P_i)\setminus\{a\}$ and $V(P_j)\setminus\{a\}$.
	\item $a$ is adjacent to at most one of $\{b_1, b_2, b_3\}$.
\end{itemize}

@@

We will give an algorithm finding pyramid, but first we need some additional definitions. If $K$ is a pyramid $(a, b_1, b_2, b_3, P_1, P_2, P_3)$ we say its \emph{frame} is the 10-tuple iff
\TODO{finish it up}

@@

First, let us enumerate all 6-tuples $b_1, b_2, b_3, s_1, s_2, s_3$ such that:
\begin{itemize}
	\item $\{b_1, b_2, b_3\}$ is a triangle
	\item for $1 \leq i < j \leq 3$, ${b_i, s_i}$ is disjoint from ${b_j, s_j}$ and $b_ib_j$ is the only edge between them
	\item there is a node $a$ adjacent to all of $s_1, s_2, s_3$ and to at most one of $b_1, b_2, b_3$, such that if $a$ is adjacent to $b_i$, then $s_i \ b_i$.
\end{itemize}

There are $O(|V(G)|^6)$ sextuples, and it takes $O(|V(G)|)$ time to check each one. For each such sextuple we follow with the rest of the algorithm.

@@

If $s_1 = b_1$ let $P_1(b_1)$ be the one-node path with node $b_1$, and let $P_1(m)$ be undefined for each $m \in M$.

If $s_1 \neq b_1$, then $P_1(b_1)$ is undefined and for all $m \in M$ we will check if all the following are true:
\begin{itemize}
	\item $m$ is nonadjacent to all of $b_2, b_3, s_2, s_3$
	\item $S_1(m)$ and $T_1(m)$ both exist
	\item $V(S_1(m) \cap T_1(m)) = \{m\}$
	\item there are no edges between $V(S_1(m) \setminus m)$ and $V(T_1(m) \setminus m)$
\end{itemize}
If so, then we assign a path $s_1-S_1(m)-m-T_1(m)-b_1$ to $P_1(m)$, otherwise we let $P_1(m)$ be undefined. It takes $O(|V(G)|^2)$ to check this, given $m$. We assign $P_2$ and $P_3$ in a similar manner. Total time of finding all $P_i(m)$ paths for a given sextuple is $O(|V(G)|^3)$.

@@

If $P_1(m_1)$ does not exist, there are no such good pairs. If it exists, color black the nodes of $M$ that either belong to $P_1(m_1)$ or have a neighbor in $P_1(m_1)$. Color all other nodes white. (We can do this in $O(|V(G)|^2)$) Then for each $m_2 \in M \cup \{b_2\}$, test whether $P_2(m_2)$ exists and contains no black nodes. We do this for all $m_1$ and get a set of all $(1,2)$-good pairs. In similar way we calculate all good $(1,3)$-pair and $(2,3)$-pairs (in $O(|V(G)|^3)$ time).

@@

% \begin{wrapfigure}[7]{r}{0.35\textwidth}
\begin{wrapfigure}{r}{0.35\textwidth}
	\centering\begin{tikzpicture}[scale=0.7,simplegraph]
	\def\ngon{5}
	\node[regular polygon,regular polygon sides=\ngon,minimum size=3cm, draw=none] (p) {};
	\foreach\x in {1,...,\ngon}{\node[] (p\x) at (p.corner \x){$v_\the\numexpr\intcalcMod{\x+3}{5}+1$};}
	%p1 - v5, p2 - v1 ...
	\draw (p1) to (p2);
	\draw (p2) to (p3);
	\draw (p3) to (p4);
	\draw (p4) to (p5);
	\draw (p1) to (p5);

  \draw[dashed] (p2) to (p5);
  \draw[dashed] (p2) to (p4);
	\draw[dashed] (p3) to (p5);

	\tikzset{decoration={snake,amplitude=.6mm,segment length=6mm,
                       post length=0mm,pre length=0mm}}
	\draw[decorate] (p2) to [out=-100, in=180] (0, -5) to [out=0, in=-80] (p5);
	\node[draw=none] at (0, -4.7) {$P$};
\end{tikzpicture}
\caption{An example of a jewel.}
\vspace{-1.5cm}
\end{wrapfigure}

Five nodes $v_1, \ldots, v_5$ and a path $P$ is a \emph{jewel} iff:

\begin{itemize}
	\item $v_1, \ldots, v_5$ are distinct nodes.
	\item $v_1v_2, v_2v_3, v_3v_4, v_4v_5, v_5v_1$ are edges.
	\item $v_1v_3, v_2v_4, v_1,v_4$ are nonedges.
	\item $P$ is a path between $v_1$ and $v_4$, such that $v_2, v_3, v_5$ have no neighbors in its inside.
\end{itemize}

@@

A configuration of type $\T_1$ is a hole of length 5. To find it we simply iterate all choices of paths of length of 4, and check if there exists a fifth node to complete the hole. See paragraph \ref{Optimizations} for more implementation details.

@@

% \begin{wrapfigure}[15]{r}{0.35\textwidth}
\begin{wrapfigure}{r}{0.35\textwidth}
	\centering\begin{tikzpicture}[scale=0.7,simplegraph]
		\node(v1) at (0,0) {$v_1$};
		\node(v2) at (1.5,0) {$v_2$};
		\node(v3) at (3,0) {$v_3$};
		\node(v4) at (4.5,0) {$v_4$};

		\draw (v1) to (v2);
		\draw (v2) to (v3);
		\draw (v3) to (v4);

		\tikzset{decoration={snake,amplitude=.6mm,segment length=6mm,
                       post length=0mm,pre length=0mm}}
		\draw[decorate] (v1) to [out=90, in=180] (4.5/2, 3) to [out=0, in=90] (v4);
		\node[draw=none] at (4.5/2, 3.5) {$P$};

		\draw[dashed] (v2) to (.7,2.5);
		\draw[dashed] (v3) to (.7,2.5);
		\draw[dashed] (v2) to (3.8,2.5);
		\draw[dashed] (v3) to (3.8,2.5);

		\node[minimum size=4mm](x1) at (1.2, -2){};
		\node[minimum size=4mm](x2) at (3.3, -2){};
		\node[draw,dotted,inner sep=3pt, circle,yscale=.5, fit={(x1) (x2)},label=below:{$X$}] {};

		\draw[dashed] (x1) to (x2);
		\draw(v1) to (x1);
		\draw(v2) to (x1);
		\draw(v4) to (x1);
		\draw(v1) to (x2);
		\draw(v2) to (x2);
		\draw(v4) to (x2);

		% \draw (a) to (b2);
\end{tikzpicture}%
\caption{An example of a $\T_2$.}%
\vspace{-1.5cm}
\end{wrapfigure}

A configuration of type $\T_2$ is a six $(v_1, v_2, v_3, v_4, P, X)$, such that:
\begin{itemize}
	\item $v_1v_2v_3v_4$ is a path in $G$.
	\item $X$ is an anticomponent of the set of all $\{v_1, v_2, v_4\}$-complete nodes.
	\item $P$ is a path in $G\setminus(X \cup \{v_2, v_3\})$ between $v_1$ and $v_4$ and no vertex in $P^*$ is $X$-complete or adjacent to $v_2$ or adjacent to $v_3$.
\end{itemize}

Checking if configuration of type $\T_2$ exists in our graph is straightforward: we enumerate all paths $v_1\ldots v_4$, calculate set of all $\{v_1, v_2, v_4\}$-complete nodes and its anticomponents. Then, for each anticomponent $X$ we check if required path $P$ exists.

To prove that existence of $\T_2$ configuration implies that the graph is not berge, we will need the Roussel-Rubio lemma \cite{RR01} as stated in \cite{MC05} that says:

\begin{lemma}[Roussel-Rubio Lemma]\label{Roussel-Rubio}
	Let $G$ be Berge, $X$ be an anticonnected subset of $V(G)$, $P$ be an odd path $p_1\ldots p_n$ in $G\setminus X$ with length at least 3, such that $p_1$ and $p_n$ are $X$-complete and $p_2, \ldots, p_{n-1}$ are not. Then:
	\begin{itemize}
		\item $P$ is of length at least 5 and there exist nonadjacent $x, y \in X$, such that there are exactly two edges between $x, y$ and $P^*$, namely $xp_2$ and $yp_{n-1}$,
		\item or $P$ is of length 3 and there is an odd antipath joining internal nodes of $P$ with interior in $X$.
	\end{itemize}
\end{lemma}

@@

\begin{lemma}
	If $G$ contains configuration of type $\T_2$ then $G$ is not Berge.
\end{lemma}
\begin{proof}
	Let $(v_1, v_2, v_3, v_4, P, X)$ be a configuration of type $\T_2$. Let us assume that $G$ is not Berge and consider the following:
	\begin{itemize}
		\item If $P$ is even, then $v_1, v_2, v_3, v_4, P, v_1$ is an odd hole,
		\item If $P$ is of length 3. \todo{I merged a couple of proofs from \cite{MC06}, check in the morning if this is correct.} Let us name its nodes $v_1, p_2, p_3, v_4$. It follows from \cref{Roussel-Rubio}, that there exists an odd antipath between $p_2$ and $p_3$ with interior in $X$. We can complete it with $v_2p_2$ and $v_2p_3$ into an odd antihole.
		\item If $P$ is odd with the length of at least 5 \todo{check in the morning}, it follows from \cref{Roussel-Rubio} that we have $x, y \in X$ with only two edges to $P$ being $xp_2$ and $yp_{n-1}$. This gives us an odd hole: $v_2, x, p_2, \ldots, p_{n-1}, y, v_2$.
	\end{itemize} 
\end{proof}

@@

A configuration of type $\T_3$ is a sequence $v_1, \ldots, v_6$, $P$, $X$, such that:
\begin{itemize}
	\item $v_1, \ldots v_6$ are distinct nodes.
	\item $v_1v_2$, $v_3v_4$, $v_1v_4$, $v_2v_3$, $v_3v_5$, $v_4v_6$ are edges, and $v_1v_3$, $v_2v_4$, $v_1v_5$, $v_2v_5$, $v_1v_6$, $v_2v_6$, $v_4v_5$ are nonedges.
	\item $X$ is an anticomponent of the set of all $\{v_1, v_2, v_5\}$-complete nodes, and $v_3$, $v_4$ are not $X$-complete.
	\item $P$ is a path of $G \setminus ( X \cup \{v_1, v_2, v_3, v_4\} )$ between $v_5$ and $v_6$ and no node in $P*$ is $X$-complete or adjacent to $v_1$ or adjacent to $v_2$.
	\item If $v_5v_6$ is an edge, then $v_6$ is not $X$-complete.
\end{itemize}

\TODO{picture?}

@@

For each triple $v_1, v_2, v_5$ of nodes such that $v_1v_2$ is an edge and $v_1v_5, v_2v_5$ are nonedges find the set $Y$ of all $\{v_1, v_2, v_5\}$-complete nodes. For each anticomponent $X$ of $Y$ find the maximal connected subset $F'$ containing $v_5$ such that $v_1, v_2$ have no neighbors in $F'$ and no node of $F'\setminus\{v_5\}$ is $X$-complete. Let $F$ be the union of $F'$ and the set of all $X$-complete nodes that have a neighbor in $F'$ and are nonadjacent to all of $v_1, v_2$ and $v_5$.

Then, for each choice of $v_4$ that is adjacent to $v_1$ and not to $v_2$ and $v_5$ and has a neighbor in $F$ (call it $v_6$) and a nonneibhbor in $X$, we test whether there is a node $v_3$, adjacent to $v_2, v_4, v_5$ and not to $v_1$, with a nonneibhbor in $X$. If there is such a node $v_3$, find $P$ -- a path from $v_6$ to $v_5$ with interior in $F'$ and return that $v_1, \ldots v_6, P, X$ is a configuration of type $\T_3$. If we exhaust our search and find none, report that graph does not contain it.

@@

We will skip the proof that each graph containing $\T_3$ is not Berge. \todo{Should we include it?} See \cite{MC05} for the proof.

@@

For a shortest odd hole $C$ in $G$, we will call a node $v \in V(G) \setminus V(C)$ $C$-\emph{major} iff the set of its neighbors in C is not contained in any 3-node path of $C$. An odd hole $C$ will be called \emph{clean} if no vertex is $C$-major.

A hole $C$ of $G$ is \emph{amenable} iff C is a shortest odd hole in $G$ of length at least 7, and for every anticonnected set $X$ of $C$-major nodes, there is an $X$-complete edge in $C$ (an edge is $X$-complete iff both its ends are $X$-complete).

The rest of the algorithm rests upon the following theorem:

\begin{theorem}
	Let $G$ be a graph, such that $G$ and $\overline{G}$ contain no Pyramid, no Jewel and no configuration of types $\T_1, \T_2$ or $\T_3$. Then every shortest hole in $G$ is amenable.
	\TODO{Any ideas on what else to say here? List all 9 steps?}
\end{theorem}

For a shortest odd hole $C$ in $G$, a subset $X$ of $V(G)$ is a \emph{near-cleaner for $C$} iff $X$ contains all $C$-major nodes and $X \cap V(C)$ is a subset of node set of some 3-node path of $C$.

\TODO{Finding and Using Half-Cleaners.}

@@

\subsubsection{Optimizations}\label{Optimizations}
Bottlenecks in performance (next path, are vectors distinct etc).\\
\TODO{In our graph preprocessing we have a pointer to next edge in order to speed up generating next path.}
\TODO{We used callgrind to get idea of methods crucial for time.}
\TODO{In general enumerating all paths is crucial. As is checking if vector has distinct values.}
\TODO{Jewels -- we iterate all possibly chordal paths and check if they are ok - much faster}
\TODO{$\T_1$ -- we iterate all paths of length 4 and check if there exists a fifth node to complete the hole - much faster than iterating nodes.}

>>>>

\documentclass{article}
\usepackage{import}

\import{./}{preamble.tex}

% Language and typesetting notes:
%
% - Color, not Colour (American English)

\author{Adrian Siwiec}
\date{\today{}}
\begin{document}

\include{titlepage}
\pagebreak

\begin{abstract}
	TODOa
\end{abstract}

@@

\section*{Definitions}
\input{definitions.tex}

@@

A definition of perfect graphs (\cref{def:perfectGraph}) states that given a graph $G$, the chromatic number and cardinality of the largest clique of \emph{its every induced subgraph} should be equal. The notion of perfect graphs was first introduced by Berge in 1961 \cite{CB61} and it indeed captures the idea of graph being ''simple'' -- in all perfect graphs the coloring problem, maximum clique problem, and maximum independent set problem can be solved in polynomial time \cite{grotschel1993}.

One of properties of perfect graphs is the \emph{perfect graph theorem} first conjured by Berge in 1961~\cite{CB61} and then proven by \Lovasz in 1972~\cite{LL72}.

\begin{theorem}[Perfect graph theorem]
	A graph is perfect if and only if its complement graph is also perfect. \todo{Should we give some proof of that here? Maybe based on proof in \cite{GC03}}
\end{theorem}

@@

Odd holes are not perfect, since their chromatic number is 3 and their largest cliques are of size 2. It is also easy to see, that an odd antihole of size $n$ has a chromatic number of $\frac{n+1}{2}$ and largest cliques of size $\frac{n-1}{2}$. A graph with no odd hole and no odd antihole is called \emph{Berge} (\cref{def:bergeGraph}) after Claude Berge who studied perfect graphs.

In 1961 Berge conjured that a graph is perfect iff it contains no odd hole and no odd antihole in what has become known as a strong perfect graph conjecture. In 2001 Chudnovsky et al. have proven it and published the proof in an over 150 pages long paper \citetitle{MC06} \cite{MC06}. The following overview of the proof will be based on this paper and on an article with the same name by Cornuéjols \cite{GC03}.

\begin{theorem}[Strong perfect graph theorem]
	A graph is perfect if and only if it is Berge.
\end{theorem}

\TODO{How long and detailed overview of the proof should we provide?}
\TODO{Znakomite pytanie. Generalnie ponieważ dowód jest drobiazgowy i trikowy, to wystarczy "z lotu ptaka" tj. to, co Chudnovsky i Seymour piszą w pracach popularnych i referują w wystąpieniach gościnnych typu te nagrania na YT. Głębiej nie ma sensu. Warto podkreślić, że technika dowodu i sam algorytm są zależne na dużo głębszym poziomie niż widać tj. ktoś mając dowód Stron Perfect Graph Theorem pewnie nie rozgryzłby algorytmu i vice versa.}

@@

@@

\begin{wrapfigure}{r}{0.35\textwidth}
	\input{tikzpictures/pyramid.tex}
	\caption{An example  of a pyramid.}%
	\vspace{-1.2cm}
	% \end{floatingfigure}
\end{wrapfigure}

A \emph{pyramid} in G is an induced subgraph formed by the union of a triangle \footnote{A triangle is a clique $K_3$.} $\{b_1,b_2,b_3\}$, three paths $\{P_1, P_2, P_3\}$ and another vertex $a$, so that:
\begin{itemize}
	\item $\forall_{1\leq i \leq 3}$ $P_i$ is a path between $a$ and $b_i$
	\item $\forall_{1\leq i < j \leq 3}$ $a$ is the only vertex in both $P_i$ and $P_j$ and $b_ib_j$ is the only edge between $V(P_i)\setminus\{a\}$ and $V(P_j)\setminus\{a\}$.
	\item $a$ is adjacent to at most one of $\{b_1, b_2, b_3\}$.
\end{itemize}

@@

@@

If $K$ is a pyramid $(a, b_1, b_2, b_3, P_1, P_2, P_3)$ we say its \emph{frame} is the 10-tuple iff
\TODO{finish it up}

First, let us enumerate all 6-tuples $b_1, b_2, b_3, s_1, s_2, s_3$ such that:
\begin{itemize}
	\item $\{b_1, b_2, b_3\}$ is a triangle
	\item for $1 \leq i < j \leq 3$, ${b_i, s_i}$ is disjoint from ${b_j, s_j}$ and $b_ib_j$ is the only edge between them
	\item there is a vertex $a$ adjacent to all of $s_1, s_2, s_3$ and to at most one of $b_1, b_2, b_3$, such that if $a$ is adjacent to $b_i$, then $s_i \ b_i$.
\end{itemize}

There are $O(|V(G)|^6)$ 6-tuples, and it takes $O(|V(G)|)$ time to check each one. For each such 6-tuple we follow with the rest of the algorithm.

@@

If $s_1 = b_1$ let $P_1(b_1)$ be the one-vertex path with vertex $b_1$, and let $P_1(m)$ be undefined for each $m \in M$.

If $s_1 \neq b_1$, then $P_1(b_1)$ is undefined and for all $m \in M$ we will check if all the following are true:
\begin{itemize}
	\item $m$ is nonadjacent to all of $b_2, b_3, s_2, s_3$
	\item $S_1(m)$ and $T_1(m)$ both exist
	\item $V(S_1(m) \cap T_1(m)) = \{m\}$
	\item there are no edges between $V(S_1(m) \setminus m)$ and $V(T_1(m) \setminus m)$
\end{itemize}
If so, then we assign a path $s_1-S_1(m)-m-T_1(m)-b_1$ to $P_1(m)$, otherwise we let $P_1(m)$ be undefined. It takes $O(|V(G)|^2)$ to check this, given $m$. We assign $P_2$ and $P_3$ in a similar manner. Total time of finding all $P_i(m)$ paths for a given 6-tuple is $O(|V(G)|^3)$.

@@

If $P_1(m_1)$ does not exist, there are no such good pairs. If it exists, color black the vertices of $M$ that either belong to $P_1(m_1)$ or have a neighbor in $P_1(m_1)$. Color all other vertices white. (We can do this in $O(|V(G)|^2)$) Then for each $m_2 \in M \cup \{b_2\}$, test whether $P_2(m_2)$ exists and contains no black vertices. We do this for all $m_1$ and get a set of all $(1,2)$-good pairs. In similar way we calculate all good $(1,3)$-pair and $(2,3)$-pairs (in $O(|V(G)|^3)$ time).

@@

% \begin{wrapfigure}[7]{r}{0.35\textwidth}
\begin{wrapfigure}{r}{0.35\textwidth}
	\input{tikzpictures/jewel.tex}
	\caption{An example of a jewel.}
	\vspace{-1.5cm}
\end{wrapfigure}

Five vertices $v_1, \ldots, v_5$ and a path $P$ form a \emph{jewel} iff:

\begin{itemize}
	\item $v_1, \ldots, v_5$ are distinct vertices.
	\item $v_1v_2, v_2v_3, v_3v_4, v_4v_5, v_5v_1$ are edges.
	\item $v_1v_3, v_2v_4, v_1,v_4$ are nonedges.
	\item $P$ is a path between $v_1$ and $v_4$, such that $v_2, v_3, v_5$ have no neighbors in its inside.
\end{itemize}

@@

A configuration of type $\T_1$ is a hole of length 5. To find it we simply iterate all choices of paths of length of 4, and check if there exists a fifth vertex to complete the hole. See paragraph \ref{Optimizations} for more implementation details.

@@

% \begin{wrapfigure}[15]{r}{0.35\textwidth}
\begin{wrapfigure}{r}{0.35\textwidth}
	\input{tikzpictures/t2.tex}
	\caption{An example of a $\T_2$.}%
	\vspace{-1.5cm}
\end{wrapfigure}

A configuration of type $\T_2$ is a tuple $(v_1, v_2, v_3, v_4, P, X)$, such that:
\begin{itemize}
	\item $v_1v_2v_3v_4$ is a path in $G$.
	\item $X$ is an anticomponent of the set of all $\{v_1, v_2, v_4\}$-complete vertices.
	\item $P$ is a path in $G\setminus(X \cup \{v_2, v_3\})$ between $v_1$ and $v_4$ and no vertex in $P^*$ is $X$-complete or adjacent to $v_2$ or adjacent to $v_3$.
\end{itemize}

Checking if configuration of type $\T_2$ exists in our graph is straightforward: we enumerate all paths $v_1\ldots v_4$, calculate set of all $\{v_1, v_2, v_4\}$-complete vertices and its anticomponents. Then, for each anticomponent $X$ we check if required path $P$ exists.

To prove that existence of $\T_2$ configuration implies that the graph is not berge, we will need the following Roussel-Rubio lemma:

\begin{lemma}[Roussel-Rubio Lemma \cite{RR01,MC05}]\label{lem:Roussel-Rubio}
	Let $G$ be Berge, $X$ be an anticonnected subset of $V(G)$, $P$ be an odd path $p_1\ldots p_n$ in $G\setminus X$ with length at least 3, such that $p_1$ and $p_n$ are $X$-complete and $p_2, \ldots, p_{n-1}$ are not. Then:
	\begin{itemize}
		\item $P$ is of length at least 5 and there exist nonadjacent $x, y \in X$, such that there are exactly two edges between $x, y$ and $P^*$, namely $xp_2$ and $yp_{n-1}$,
		\item or $P$ is of length 3 and there is an odd antipath joining internal vertices of $P$ with interior in $X$.
	\end{itemize}
	\TODO{We may use this lemma quite often, might want to provide proof if so.}
\end{lemma}

@@

\begin{lemma}
	If $G$ contains configuration of type $\T_2$ then $G$ is not Berge.
\end{lemma}
\begin{proof}
	Let $(v_1, v_2, v_3, v_4, P, X)$ be a configuration of type $\T_2$. Let us assume that $G$ is not Berge and consider the following:
	\begin{itemize}
		\item If $P$ is even, then $v_1, v_2, v_3, v_4, P, v_1$ is an odd hole,
		\item If $P$ is of length 3. \todo{I merged a couple of proofs from \cite{MC06}, check in the morning if this is correct.} Let us name its vertices $v_1, p_2, p_3, v_4$. It follows from \cref{lem:Roussel-Rubio}, that there exists an odd antipath between $p_2$ and $p_3$ with interior in $X$. We can complete it with $v_2p_2$ and $v_2p_3$ into an odd antihole.
		\item If $P$ is odd with the length of at least 5 \todo{check in the morning}, it follows from \cref{lem:Roussel-Rubio} that we have $x, y \in X$ with only two edges to $P$ being $xp_2$ and $yp_{n-1}$. This gives us an odd hole: $v_2, x, p_2, \ldots, p_{n-1}, y, v_2$.
	\end{itemize}
\end{proof}

@@

A configuration of type $\T_3$ is a sequence $v_1, \ldots, v_6$, $P$, $X$, such that:
\begin{itemize}
	\item $v_1, \ldots v_6$ are distinct vertices.
	\item $v_1v_2$, $v_3v_4$, $v_1v_4$, $v_2v_3$, $v_3v_5$, $v_4v_6$ are edges, and $v_1v_3$, $v_2v_4$, $v_1v_5$, $v_2v_5$, $v_1v_6$, $v_2v_6$, $v_4v_5$ are nonedges.
	\item $X$ is an anticomponent of the set of all $\{v_1, v_2, v_5\}$-complete vertices, and $v_3$, $v_4$ are not $X$-complete.
	\item $P$ is a path of $G \setminus ( X \cup \{v_1, v_2, v_3, v_4\} )$ between $v_5$ and $v_6$ and no vertex in $P*$ is $X$-complete or adjacent to $v_1$ or adjacent to $v_2$.
	\item If $v_5v_6$ is an edge, then $v_6$ is not $X$-complete.
\end{itemize}

\TODO{picture!}

@@

For each triple $v_1, v_2, v_5$ of vertices such that $v_1v_2$ is an edge and $v_1v_5, v_2v_5$ are nonedges find the set $Y$ of all $\{v_1, v_2, v_5\}$-complete vertices. For each anticomponent $X$ of $Y$ find the maximal connected subset $F'$ containing $v_5$ such that $v_1, v_2$ have no neighbors in $F'$ and no vertex of $F'\setminus\{v_5\}$ is $X$-complete. Let $F$ be the union of $F'$ and the set of all $X$-complete vertices that have a neighbor in $F'$ and are nonadjacent to all of $v_1, v_2$ and $v_5$.

Then, for each choice of $v_4$ that is adjacent to $v_1$ and not to $v_2$ and $v_5$ and has a neighbor in $F$ (call it $v_6$) and a nonneibhbor in $X$, we test whether there is a vertex $v_3$, adjacent to $v_2, v_4, v_5$ and not to $v_1$, with a nonneibhbor in $X$. If there is such a vertex $v_3$, find $P$ -- a path from $v_6$ to $v_5$ with interior in $F'$ and return that $v_1, \ldots v_6, P, X$ is a configuration of type $\T_3$. If we exhaust our search and find none, report that graph does not contain it.

@@

We will skip the proof that each graph containing $\T_3$ is not Berge. See paragraph 6.7 of \cite{MC05} for the proof.

@@

\begin{theorem}
	Let $G$ be a graph, such that $G$ and $\overline{G}$ contain no Pyramid, no Jewel and no configuration of types $\T_1, \T_2$ or $\T_3$. Then every shortest hole in $G$ is amenable.
	\TODO{Any ideas on what else to say here? List all 9 steps?}
	\TODO{Dobre pytanie. Te 9 kroków to da się zawrzeć w 1-2 zdaniach każde? Jeśli nie, to tylko napisałbym ogólnikowo, że "jest sekwencja kroków, która to robi, tu jest odnośnik"}
\end{theorem}

\TODO{Finding and Using near-cleaners.}

@@

\subsubsection{Optimizations}\label{Optimizations}
Bottlenecks in performance (next path, are vectors distinct etc).\\
\TODO{In our graph preprocessing we have a pointer to next edge in order to speed up generating next path.}
\TODO{We used callgrind to get idea of methods crucial for time.}
\TODO{In general enumerating all paths is crucial. As is checking if vector has distinct values.}
\TODO{Jewels -- we iterate all possibly chordal paths and check if they are ok - much faster}
\TODO{$\T_1$ -- we iterate all paths of length 4 and check if there exists a fifth vertex to complete the hole - much faster than iterating vertices.}
